---
title: "Wilshire 5000 ARMA GARCH Models"
output: html_notebook
---

```{r}
rm(list = ls())

library("AER")
library("ggplot2")
library("sandwich")
library("stargazer")
library("survival")
library("plm")
library("margins")
library("readxl")
library("xts")
library("dynlm")
library("zoo")
library("urca")
library("strucchange")
library("orcutt")
library("fGarch")
library("quantmod")
library("tidyr")

setwd("~/Econometrics/Testing SP500")

```

Loading Data from the Wilshire 5000 Stock Price Index Data

```{r}

w5000 <- read_excel(path = "W5000.xlsx")
w5000
```
Transforming Columns
```{r}

w5000$Date <- as.Date(w5000$observation_date)
w5000$WILL5000INDFC <- as.numeric(w5000$WILL5000INDFC)
w5000 <- na.omit(w5000)

```


Computing Daily Percentage Changes using the Delt function from the package quantmod

```{r}
N <- length(w5000$WILL5000INDFC)
w5000_pc_xts <- xts(100*Delt(w5000$WILL5000INDFC)[-1], w5000$Date[-N])

```

Setting up a list of dates to convert data into xts objects

```{r}
dates <- w5000$Date[-N]

```


```{r}
#Plotting Percentage Changes
plot(as.zoo(w5000_pc_xts),
     ylab = "Percent", 
     xlab = "Date", 
     main = "Daily Percentage Changes \n Wilshire 5000 Stock Price Index", 
     type = "l", 
     col = "steelblue", 
     lwd = 0.5,
)

#Adding horizontal line at y = 0


```

Fitting a model to our data, we specify the type of ARMA model to use to model 
the values of the outcome and the type of GARCH to model conditional heteroskedasticity.

Using ARMA (1,1) GARCH(1,0) model

```{r}
GARCH_Wilshire <- garchFit(~ arma(1,1) + garch(1,0), data = w5000_pc_xts, trace = F)
summary(GARCH_Wilshire)


```

Considering model's predictions of actual value of percentage change

```{r}
w5000_pc_fitted_xts <- xts(GARCH_Wilshire@fitted, dates)

#to view whats happening, narrowing in on one small time period: 2008-2009
w5000_pc_2008to2009_xts <- w5000_pc_xts["2008::2009"]
w5000_pc_fitted_2008to2009_xts <- w5000_pc_fitted_xts["2008::2009"]

```

Plotting percentage changes and model fitted values

```{r}
plot(as.zoo(w5000_pc_2008to2009_xts),
     type = "l", 
     col = "steelblue", 
     ylab = "Percent", 
     xlab = "Date", 
     main = "ARMA(1,1), GARCH(1,0) \n Predicted Percent Change", 
     lwd = 0.25)

#adding most fitted values of percentage change to plot
lines(as.zoo(w5000_pc_fitted_2008to2009_xts), 
      col = "forestgreen", 
      lwd = 1.5) ####Rework code

legend("topright",
       lty = c(1, 1),
       lwd = c(0.2, 1),
       cex = 0.8,
       col = c("steelblue", "forestgreen"),
       legend = c("Actual Percentage Change", "Predicted Percentage Change"))

```

Computing deviations of percentage changes from their mean

```{r}
demean_w5000_pc <- w5000_pc_xts - GARCH_Wilshire@fit$coef[1]

```

The demean_w5000_pc_pcsd time series indicates the mean estimated by our model, 
plus the conditional standard error: this is our model fit of the 
conditional standard error at a given time, based on other values

```{r}
demean_w5000_pc_pcsd <- GARCH_Wilshire@fit$coef[1] + GARCH_Wilshire@sigma.t

```

The demean_w5000_pc_mcsd time series indicates the mean estimated by our model, 
minus the conditional standard error: this is our model fit of the 
conditional standard error at a given time, based on other values

```{r}
demean_w5000_pc_mcsd <- GARCH_Wilshire@fit$coef[1] - GARCH_Wilshire@sigma.t

```

Together, the plus/minus one conditional standard error gives a confidence 
interval for the predicted volatility of the time series.

Converting these into xts objects:

```{r}
demean_w5000_pc_xts <- xts(demean_w5000_pc, dates)
demean_w5000_pc_pcsd_xts <- xts(demean_w5000_pc_pcsd, dates)
demean_w5000_pc_mcsd_xts <- xts(demean_w5000_pc_mcsd, dates)

```

Focusing on small time period: 2008 to 2009

```{r}
demean_w5000_pc_2008to2009 <- demean_w5000_pc_xts["2008::2009"]
demean_w5000_pc_pcsd_2008to2009 <- demean_w5000_pc_pcsd_xts["2008::2009"]
demean_w5000_pc_mcsd_2008to2009 <- demean_w5000_pc_mcsd_xts["2008::2009"]

```

Plotting deviation of percentage changes from mean

```{r}
plot(as.zoo(demean_w5000_pc_2008to2009), 
     type = "l", 
     col = "steelblue",
     ylab = "Percent", 
     xlab = "Date",
     main = "ARMA(1,1)-GARCH(1,0) \n Estimated Bands of +- One Conditional Standard Deviation",
     lwd = 0.2,
     )
# add horizontal line at y = 0
#abline(0, 0)
# add GARCH confidence bands (one standard deviation) to the plot
lines(as.zoo(demean_w5000_pc_pcsd_2008to2009), 
      col = "darkred", 
      lwd = 0.5)
lines(as.zoo(demean_w5000_PC_mcsd_2008to2009), 
      col = "yellow", 
      lwd = 0.5)
legend("topright",
       lty = c(1, 1),
       lwd = c(0.5, 0.5),
       cex = 0.7,
       col = c("steelblue", "darkred"),
       legend = c("Actual Percentage Change", "+/- Conditional Standard Deviation"))
```

Fitting a new model to the data specifying the type of ARMA model to be used and
the GARCH model to be used to model conditional heteroskedasticity in the data

```{r}

GARCH_Wilshire_2 <- garchFit(~ arma(1,1) + garch(1,1), data = w5000_pc_xts, trace = F)
summary(GARCH_Wilshire_2)

```

All coefficients under the error analysis are statistically significant at the 0.0001 level of significance.

Predicting the actual value of percentage change:

```{r}
w5000_pc_fitted_xts_2 = xts(GARCH_Wilshire_2@fitted, dates)

#focusing on the period 2008 to 2009

w5000_pc_xts_2008to2009_2 <- w5000_pc_xts_2["2008::2009"]

w5000_pc_fitted_xts_2008to2009_2 <- w5000_pc_fitted_xts_2["2008::2009"]
```

Plotting percentage changes 

```{r}

plot(as.zoo(w5000_pc_xts_2008to2009_2), 
      type = "l", 
      col = "steelblue",
      ylab = "Percent", 
      xlab = "Date",
      main = "ARMA(1,1)-GARCH(1,1) \n Predicted Percentage Change",
      lwd = 0.25)

# add horizontal line at y = 0
# abline(0, 0)
# # add the model-fitted values of the percentage change to the plot:
lines(as.zoo(w5000_pc_fitted_xts_2008to2009_2), 
       col = "forestgreen", 
       lwd = 1.5)
legend("topright",
        lty = c(1, 1),
        lwd = c(0.2, 1),
        cex = 0.8,
        col = c("steelblue", "forestgreen"),
        legend = c("Actual Percentage Change", "Predicted Percentage Change"))


```

Computing deviations of percentage changes from their mean:

```{r}
demean_w5000_2 <- w5000_pc_xts_2 - GARCH_Wilshire_2@fit$coef[1]


```

Also, estimating for mean and conditional standard error:

The demean_W5000_pc_pcsd_2 time series indicates the mean estimated by our model, 
plus the conditional standard error: this is our model fit of the 
conditional standard error at a given time, based on other values.

The demean_W5000_PC_Mcsd time series indicates the mean estimated by our model, 
minus the conditional standard error: this is our model fit of the 
conditional standard error at a given time, based on other values.

Together, the plus/minus one conditional standard error gives a confidence 
interval for the predicted volatility of the time series.

```{r}
demean_w5000_pc_pcsd_2 <- GARCH_Wilshire_2@fit$coef[1] + GARCH_Wilshire_2@sigma.t
demean_w5000_pc_mcsd_2 <- GARCH_Wilshire_2@fit$coef[1] - GARCH_Wilshire_2@sigma.t
```

Converting these into xts objects:

```{r}
demean_w5000_xts_2 <- xts(demean_w5000_2, dates) 
demean_w5000_pc_pcsd_xts_2 <- xts(demean_w5000_pc_pcsd_2, dates)
demean_w5000_pc_mcsd_xts_2 <- xts(demean_w5000_pc_mcsd_2, dates)

```

Focusing on periods 2008 to 2009:

```{r}
demean_w5000_xts_2008to2009_2 <- demean_w5000_xts_2["2008::2009"]
demean_w5000_pc_pcsd_xts_2008to2009_2 <- demean_w5000_pc_pcsd_xts_2["2008::2009"]
demean_w5000_pc_mcsd_xts_2008to2009_2 <- demean_w5000_pc_mcsd_xts_2["2008::2009"]
```

Plotting for period 2008 to 2009:

```{r}
plot(as.zoo(demean_w5000_xts_2008to2009_2), 
     type =  "l", 
     col = "steelblue", 
     ylab = "Percent", 
     xlab = "Date", 
     main = "ARMA(1,1), GARCH(1,1) \n Estimated Bands +- One Conditional Standard Deviation", 
     lwd = 0.2)
#abline(0, 0)
lines(as.zoo(demean_w5000_pc_pcsd_xts_2008to2009_2), 
      col = "darkred", 
      lwd = 0.2)
lines(as.zoo(demean_w5000_pc_mcsd_xts_2008to2009_2), 
      col = "darkred", 
      lwd = 0.2)
legend("topright",
       lty = c(1, 1),
       lwd = c(0.5, 0.5),
       cex = 0.7,
       col = c("steelblue", "darkred"),
       legend = c("Actual Percentage Change", "+/- Conditional Standard Deviation"))
```
Fitting an ARMA(3,3) to our model and GARCH(3,3) to for conditional heteroskedasticity 

```{r}
GARCH_Wilshire_3 <- garchFit(~ arma(3,3) + garch(3,3), data = w5000_pc_xts, trace = F)
summary(GARCH_Wilshire_3)


```

Considering model predictions of the actual value of the percentage change 
and focusing on the period of 2008 to 2009:

```{r}
w5000_pc_fitted_xts_3 <- xts(GARCH_Wilshire_3@fitted, dates)

#focusing on 2008 to 2009

w5000_pc_xts_2008to2009_3 <- w5000_pc_xts["2008::2009"]
w5000_pc_fitted_xts_2008to2009_3 <- w5000_pc_fitted_xts_3["2008::2009"] 

```

Plotting percent changes and model fitted values:

```{r}
plot(as.zoo(w5000_pc_xts_2008to2009_3), 
     type = "l", 
     col = "steelblue", 
     ylab = "Percent", 
     xlab = "Date", 
     main = "ARMA(3,3)-GARCH (3,3) \n Predicted Percentage Change", 
     lwd = 0.25)
abline(0,0)
lines(as.zoo(w5000_pc_fitted_xts_2008to2009_3), 
      col = "forestgree", 
      lwd = 0.25)


```

Computing deviations from mean and conditional standard deviations

```{r}
demean_w5000_3 <- w5000_pc_xts - GARCH_Wilshire_3@fit$coef[1]
demean_w5000_pc_pcsd_3 <- GARCH_Wilshire_3@fit$coef[1] + GARCH_Wilshire_3@sigma.t
demean_w5000_pc_mcsd_3 <- GARCH_Wilshire_3@fit$coef[1] - GARCH_Wilshire_3@sigma.t

```

Making xts objects

```{r}
demean_w5000_3_xts <- xts(demean_w5000_3, dates)
demean_w5000_pc_pcsd_3_xts <- xts(demean_w5000_pc_pcsd_3, dates)
demean_w5000_pc_mcsd_3_xts <- xts(demean_w5000_pc_mcsd_3, dates)
```

Focusing on the period of 2008 to 2009:

```{r}
demean_w5000_3_2008to2009_xts <- demean_w5000_3_xts["2008::2009"]
demean_w5000_pc_pcsd_3_2008to2009_xts <- demean_w5000_pc_pcsd_3_xts["2008::2009"]
demean_w5000_pc_mcsd_3_2008to2009_xts <- demean_w5000_pc_mcsd_3_xts["2008::2009"]

```

Plotting percentage changes from mean:

```{r}
plot(as.zoo(demean_w5000_3_2008to2009_xts), 
     type = "l", 
     col = "steelblue", 
     ylab = "Percent", 
     xlab = "Date", 
     main = "ARMA(3,3)-GARCH(3,3) \n Estimated BAnds of +- One Conditional Standard Deviation", 
     lwd = 0.25)
abline(0,0)
lines(as.zoo(demean_w5000_pc_pcsd_3_2008to2009_xts), 
      col = "darkred", 
      lwd = 0.5)
lines(as.zoo(demean_w5000_pc_mcsd_3_2008to2009_xts), 
      col = "darkred", 
      lwd = 0.5)
legend("topright",
       lty = c(1, 1),
       lwd = c(0.5, 0.5),
       cex = 0.7,
       col = c("steelblue", "darkred"),
       legend = c("Actual Percentage Change", "+/- Conditional Standard Deviation"))

```










